---
title: "PSYR 6003 - Assignment 2"
format: html
editor: visual
---

Install and library all required packages

```{r}
install.packages("tidyverse")
install.packages("haven")
install.packages("flexplot")
install.packages("dplyr")
install.packages("apaTables")
install.packages("MASS")
install.packages("patchwork")

library(tidyverse)
library(haven)
library(flexplot)
library(dplyr)
library(apaTables)
library(MASS)
library(patchwork)
```

Read in Data set

```{r}
newData <- read_sav("~/STUPIDFOLDER/PSYR6003.A2.sav")
View(newData)
```

Clean data and prepare for analysis

```{r}
#change sex to a factor 
goodData <- newData %>%
  drop_na() %>% 
  mutate(sex = as.factor(sex)) %>% 
  mutate(sex = recode_factor(sex, Female = "0", Male = "1", "Other (specify)" = "2"))

#Reverse coding tipm.CONS2.3y
greatData <- goodData %>% 
  mutate(across(matches("tipm.CONS2.3y"), ~ 8 - .))
view(goodData)

##Means Subscales
excellentData <- greatData %>% 
  mutate(mps.mean = rowMeans(across(starts_with("mps")), na.rm = F),
         negative.mean = rowMeans(across(guilt1.3y:host3.3y), na.rm = F),
         tipm.mean = rowMeans(across(starts_with("tipm")), na.rm = F)
         )
view(excellentData)
```

Perform univariate analysis using the GLM

```{r}
##Visualize relationships
a= flexplot(negative.mean ~ 1, data = excellentData)
b=flexplot(tipm.mean ~ 1, data = excellentData)
c= flexplot(mps.mean~ 1, data = excellentData)
d= flexplot(sex~1, data = excellentData)
plot1<- a+b+c+d
##ggsave(filename = "flexplot_Univariate_relationships.jpeg",plot1)

#Step 1: Fit the model
NegativeAffect = lm(negative.mean ~ 1, excellentData)
#Step 1.5: Summarize
summary(NegativeAffect)
#Step 2: Visualize the data
visualize(NegativeAffect)
#Step 3: Compute Estimates
estimates(NegativeAffect)

#Step 1: Fit the model
Conscientiousness = lm(tipm.mean ~ 1, excellentData)
#Step 1.5: Summarize
summary(Conscientiousness)
#Step 2: Visualize the data
visualize(Conscientiousness)
#Step 3: Compute Estimates
estimates(Conscientiousness)

#Step 1: Fit the model
Perfectionism = lm(mps.mean ~ 1, excellentData)
#Step 1.5: Summarize
summary(Perfectionism)
#Step 2: Visualize the data
visualize(Perfectionism)
#Step 3: Compute Estimates
estimates(Perfectionism)
```

Test Hypothesis 1 - Sex, conscientiousness, and SPP will all significantly predict negative affect. Women will show more negative affect compared to men, and conscientiousness will be negatively related to negative affect.

Full model: Negative Affect = b0 + b1 \* sex + b2 \* Conscientiousness + b3 \* Perfectionsism + e

Reduced model : Negative Affect = b0 + e

```{r}
## Hypothesis 1 GLM
##Visualize relationships
Fig1 <- flexplot(negative.mean~tipm.mean | sex+ mps.mean, data = excellentData, method="lm")
Fig2 <- flexplot (negative.mean ~ mps.mean | sex + tipm.mean, data = excellentData, method = "lm")
H1 <-Fig1 + Fig2
##ggsave(filename = "flexplot_multivariate_relationships.jpeg", H1 ) ##Save 

#Step 1: Fit the model
model1 <- lm(negative.mean ~ sex + tipm.mean + mps.mean, data = excellentData)
# Step 2: Visualize
visualize(model1)
#Step 3: Compute Estimates
estimates(model1)
```

Sensitivity Analysis + Considerations

```{r}
## Is having a participant 'Other' an outlier? Does it need to be removed? Lets find out.
seperatedData = excellentData %>% 
  filter(sex != "2")
##Visualize relationships
Fig1 <- flexplot(negative.mean~tipm.mean | sex+ mps.mean, data = seperatedData, method="lm")
Fig2 <- flexplot (negative.mean ~ mps.mean | sex + tipm.mean, data = seperatedData, method = "lm")
H1 <-Fig1 + Fig2
ggsave(filename = "flexplot_multivariate_relationshipsNoOther.jpeg", H1 )
#Step 1: Fit the model
model2 <- lm(negative.mean ~ sex + tipm.mean + mps.mean, data = seperatedData)
# Step 2: Visualize
visualize(model2)
ggsave(filename = "model2NoOther.jpeg", model2)
#Step 3: Compute Estimates
estimates(model2)

## Male outlier
stupidData= excellentData %>% 
  filter(sex ==1)
menOnly<-flexplot(negative.mean ~ tipm.mean, data = stupidData, method = "lm")
ggsave(filename = "menOutlier.jpeg",menOnly )

##IS this a quadratic?
model3 = lm(negative.mean ~ sex + tipm.mean + mps.mean + I(tipm.mean^2), data = excellentData)
model4 = lm(negative.mean ~ sex + tipm.mean + mps.mean + I(mps.mean^2), data = excellentData)
visualize(model3)+visualize(model4)

##apa.aov.table(model3,filename = "model3.doc")
model.comparison(model1, model3)
model.comparison(model1, model4)
```

```{r}
##Compare models based on sex and other outliers
model.comparison(model1, model2)
```

```{r}
## What model should we use? Compare full VS reduced model
reducedModel = lm(negative.mean ~ 1, data = excellentData)
model.comparison(model1, reducedModel)
estimates(model1)
```

Test Hypothesis 2 - SPP will predict unique variance in negative affect over and above sex and conscientiousness in a meaningful way.

Full model: Negative Affect = b0 + b1\*Perfectionism + b2\*Conscientiousness + b3\*sex + e Reduced model: Negative Affect = b0 + b1\*sex + b2\*Conscientiousness + e

```{r}
##Hypothesis 2 GLM
##Visualize model
flexplot(negative.mean ~ sex +tipm.mean | mps.mean, data = excellentData)

## Visualize connections between all variables
added.plot(negative.mean ~ mps.mean, data = excellentData, method= "lm", lm_formula = negative.mean ~ sex + tipm.mean)
```

```{r}
##Visualize relationships
flexplot(negative.mean ~sex + tipm.mean, data = excellentData)

##Create Model
H2reducedModel = lm(negative.mean ~ sex + tipm.mean, data = excellentData)

##Visualize and compare against full model
visualize(H2reducedModel) + visualize(model1)
##Compute estimates
estimates(H2reducedModel)

##compare models to assess variance
model.comparison(model1, H2reducedModel)
```

Create Tables

```{r}

## Create tables for analysis + results
excellentDataTable <- apa.cor.table(excellentData,
                                    table.number = 1,
                                    filename = "dataSetTable.doc")
##Regression Table
model1_table <- apa.reg.table(model1, 
                              table.number = 2, 
                              filename = "model1Table.doc")
modelH2_table <- apa.reg.table(H2reducedModel,
                               table.number = 3,
                               filename = "H2reducedmodel.doc")
model1_anovaTable <- apa.aov.table(model1,
                                 table.number = 3,
                                  filename = "model1_anovaTable.doc")
```
