---
title: "Assignment 4"
format: html
editor: visual
---

## Install and Library packages

```{r}
install.packages("tidyverse")
install.packages("haven")
install.packages("flexplot")
install.packages("dplyr")
install.packages("apaTables")
install.packages("MASS")
install.packages("patchwork")
install.packages("nlme")
install.packages("lme4")
install.packages("Matrix")

library(tidyverse)
library(haven)
library(flexplot)
library(dplyr)
library(apaTables)
library(MASS)
library(patchwork)
library(nlme)
library(lme4)
library(Matrix)

```

```{r}
newData <- read_sav("~/STUPIDFOLDER/P6003.A4.sav")
View(newData)
```

```{r}
goodData <- newData %>% 
  na.omit
view(goodData)
## data looks good! MOving onto to next steps
```

Test a model where extroversion and neuroticism are predictors and the dependent variable is life satisfaction.

H1 - Extroversion will be positively associated with life satisfaction.

H2 - Neuroticism will be negatively associated with life satisfaction.

H3 - The effects will be similar for both levels 1 (within participants over time) and level 2 (between participants).

```{r}
##Visualize relationships
a= flexplot(swl ~ 1, data = goodData)
b=flexplot(tipm.E ~ 1, data = goodData)
c= flexplot(tipm.N~ 1, data = goodData)
plot1<-a+b+c
##ggsave("univariate_relationships.jpeg", plot1)

##Baseline model
## Step 1: Fit Baseline Model
##gives us total variability!
reducedModel<- lmer(swl ~1+(1|id), data=goodData)
summary(reducedModel)

## Step 2: Calculate ICC
##higher the value = less independant the data is
icc(reducedModel)

## Step 3: Create Fixed and Random models

## Fixed effects Model
fixedModel <- lmer(swl~tipm.N+tipm.E+(1|id), goodData)
summary(fixedModel)

model.comparison(reducedModel,fixedModel)

##Random Mixed Effects Model
randomModel<-lmer(swl~tipm.N+tipm.E+(tipm.N+tipm.E|id), goodData)
summary(randomModel)

##Step 4: Model Comparison

## Fixed extroversion effect. 
fixed_extraversion <- lmer(swl ~ tipm.E + (1|id), data = goodData)
summary(fixed_extraversion)

#Random + fixed extroversion effect. This model is saying there is an average effect of Extraversion on Life Satistafaction, but that changes across people
random_Extraversion <- lmer (swl ~ tipm.E + (tipm.E|id), data = goodData)
summary(random_Extraversion)

#Do a model comparison
model.comparison(fixed_extraversion, random_Extraversion)

#Take the best fitting model and add the next variable you theoretically want to test, in this case it's our predictor neurotisicm added as a fixed effect only.
fixed_neuroticsm <- lmer (swl ~ tipm.E + tipm.N + (tipm.E|id), data = goodData)
summary(fixed_neuroticsm)
model.comparison(random_Extraversion, fixed_neuroticsm)

#Add the final random neurotisicm variable to test it
random_Neurotiscm <- lmer (swl ~ tipm.E + tipm.N + (tipm.E + tipm.N|id), data = goodData)


model.comparison(fixed_neuroticsm, random_Neurotiscm)
model.comparison(fixedModel, random_Neurotiscm)
summary(random_Neurotiscm)
```

```{r}

#Visualize the model 
visualize(random_Neurotiscm, plot = "model")
#Get diagnostics
visualize(random_Neurotiscm, plot = "residuals" )

#Get summary
summary(random_Neurotiscm)

#Get key estimates.
## This will NOT work sorry
detach("package:lmerTest", unload = TRUE)
estimates(random_Neurotiscm)

#Get p-values
library(lmerTest)

#Get traditional R2 estimates
#install.packages("performance")
library(performance)

r2(random_Neurotiscm)

##Generating table with CI for Models
install.packages("sjPlot")
library(sjPlot)
tab_model(random_Neurotiscm)
```
